---
title: "R Notebook"
author: "Ryan Peckner"
date: "08/04/2021"
output:
  html_document:
    df_print: paged
---

```{r, cache=TRUE}

delta = 0.1
alpha1 = 0.3
s = 0.22

source("CCE_Lab1_MainFunctions.R")
source("CCE_Lab3_MainFunctions.R")
forcings2 = read.csv("CCE_Lab2_Outputs.csv",stringsAsFactors = FALSE)
forcings3 = read.csv("CCE_Lab3_Outputs.csv",stringsAsFactors = FALSE)
forcings4 = read.csv("CCE_Lab4_Outputs.csv",stringsAsFactors = FALSE)

Psi = 0.003
Damage = function(temperature) {
  return(Psi*(temperature^2))
}

options(digits=15)

forcings6 = forcings4
forcings6$temp.increase_bau = numeric(nrow(forcings6))
forcings6$temp.increase_bau[1] = forcings2$temp.increase.bau[1]
forcings6$capital.stock_bau = numeric(nrow(forcings6))
forcings6$capital.stock_bau[1] = 139.65
forcings6$population = forcings2$population
forcings6$population.millions = 1e-6*forcings6$population
forcings6$gross.output_bau = numeric(nrow(forcings6))
forcings6$kaya.emissions_bau = numeric(nrow(forcings6))
forcings6$total.emissions_bau = numeric(nrow(forcings6))
forcings6$damages_bau = numeric(nrow(forcings6))

for (t in 1:nrow(forcings6)) {
  capital = CapitalStock(t,abatement_cost = rep(0,nrow(forcings6)),
                         total_factor_productivity = forcings3$total.factor.productivity,
                         population = forcings6$population.millions)
  forcings6$capital.stock_bau[t] = capital
  gross_output = GrossOutput(t,total_factor_productivity = forcings3$total.factor.productivity,
                       capital = rep(capital,nrow(forcings6)),
                       population = forcings6$population.millions)
  forcings6$gross.output_bau[t] = gross_output
  kaya = KayaEmissionsFromGDP(t,population = forcings6$population,
                              output = forcings6$gross.output_bau[t],
                              energy_intensity = forcings2$energy.intensity,
                              emission_intensity = forcings2$carbon.intensity)
  forcings6$kaya.emissions_bau[t] = kaya
  forcings6$total.emissions_bau[t] = TotalEmissions(t,forcings6$kaya.emissions_bau,
                                                 forcings2$land.use.emissions)
    net_output = NetOutput(t,abatement_cost = rep(0,nrow(forcings6)),
                         damages = forcings6$damages_bau,
                         gross_output = forcings6$gross.output_bau)
  forcings6$net.output_bau[t] = net_output
  forcings6$temp.increase_bau[t] = GlobalAvgTempIncrease_FromEmissions(t,forcings6$total.emissions_bau)
  forcings6$damages_bau[t] = Damage(forcings6$temp.increase_bau[t])
}



```

Since Kaya emissions are measured in MtC, we will represent an additional ton of carbon by adding 1e-6 to the total emissions in 2011:

```{r, cache=TRUE}

forcings6$temp.increase_additional.ton.carbon = numeric(nrow(forcings6))
forcings6$temp.increase_additional.ton.carbon[1] = forcings2$temp.increase.bau[1]
forcings6$capital.stock_additional.ton.carbon = numeric(nrow(forcings6))
forcings6$capital.stock_additional.ton.carbon[1] = 139.65
forcings6$population = forcings2$population
forcings6$population.millions = 1e-6*forcings6$population
forcings6$gross.output_additional.ton.carbon = numeric(nrow(forcings6))
forcings6$kaya.emissions_additional.ton.carbon = numeric(nrow(forcings6))
forcings6$total.emissions_additional.ton.carbon = numeric(nrow(forcings6))
forcings6$damages_additional.ton.carbon = numeric(nrow(forcings6))
forcings6$total.emissions_additional.ton.carbon = numeric(nrow(forcings6))
forcings6$land.use.emissions_additional.ton.carbon = forcings2$land.use.emissions

for (t in 1:nrow(forcings6)) {
  capital = CapitalStock(t,abatement_cost = rep(0,nrow(forcings6)),
                         total_factor_productivity = forcings3$total.factor.productivity,
                         population = forcings6$population.millions)
  forcings6$capital.stock_additional.ton.carbon[t] = capital
  gross_output = GrossOutput(t,total_factor_productivity = forcings3$total.factor.productivity,
                       capital = rep(capital,nrow(forcings6)),
                       population = forcings6$population.millions)
  forcings6$gross.output_additional.ton.carbon[t] = gross_output
  kaya = KayaEmissionsFromGDP(t,population = forcings6$population,
                              output = forcings6$gross.output_additional.ton.carbon[t],
                              energy_intensity = forcings2$energy.intensity,
                              emission_intensity = forcings2$carbon.intensity)
  forcings6$kaya.emissions_additional.ton.carbon[t] = kaya
  forcings6$total.emissions_additional.ton.carbon[t] =  
                TotalEmissions(t,forcings6$kaya.emissions_additional.ton.carbon,
                                  forcings6$land.use.emissions_additional.ton.carbon)
  if (t==2) forcings6$total.emissions_additional.ton.carbon[t] = forcings6$total.emissions_additional.ton.carbon[t] + 1e-6
  net_output = NetOutput(t,abatement_cost = rep(0,nrow(forcings6)),
                         damages = forcings6$damages_additional.ton.carbon,
                         gross_output = forcings6$gross.output_additional.ton.carbon)
  forcings6$net.output_additional.ton.carbon[t] = net_output
  forcings6$temp.increase_additional.ton.carbon[t] = GlobalAvgTempIncrease_FromEmissions(t,forcings6$total.emissions_additional.ton.carbon)
  forcings6$damages_additional.ton.carbon[t] = Damage(forcings6$temp.increase_additional.ton.carbon[t])

}

    

```

Now we will compute the marginal damages. First, we want to convert damages from share of gross output to actual dollar value, which we acheive by multiplying the computed damages with gross output and then converting from trillions of dollars to dollars alone:

```{r}

forcings6$damages_bau_dollars = 
        1e12*forcings6$damages_bau*forcings6$gross.output_bau
forcings6$damages_additional.ton.carbon_dollars = 
        1e12*forcings6$damages_additional.ton.carbon*forcings6$gross.output_additional.ton.carbon
forcings6$marginal.damages_additional.ton.carbon_dollars = 
  forcings6$damages_additional.ton.carbon_dollars - forcings6$damages_bau_dollars

```

These marginal damages are computed per ton of carbon, and we will also compute the marginal damages per ton of CO2:

```{r}

forcings6$marginal.damages_additional.ton.CO2_dollars = 
  (12/44)*forcings6$marginal.damages_additional.ton.carbon_dollars 

```

Now we introduce a function to compute the net present value of the marginal damages with discount rate as a parameter.

```{r}

DiscountFactor = function(t,discount_rate) {
  return(1/((1+discount_rate)^t))
}

forcings6$discount.factor_0.03discount.rate = 
  sapply(seq(0,nrow(forcings6)-1),DiscountFactor,discount_rate = 0.03)
forcings6$discount.factor_0.025discount.rate = 
  sapply(seq(0,nrow(forcings6)-1),DiscountFactor,discount_rate = 0.025)
forcings6$discount.factor_0.05discount.rate = 
  sapply(seq(0,nrow(forcings6)-1),DiscountFactor,discount_rate = 0.05)

forcings6$net.present.value_marginal.damages_additional.ton.carbon_0.03discount.rate = 
     forcings6$marginal.damages_additional.ton.carbon_dollars*forcings6$discount.factor_0.03discount.rate
forcings6$net.present.value_marginal.damages_additional.ton.CO2_0.03discount.rate = 
     forcings6$marginal.damages_additional.ton.CO2_dollars*forcings6$discount.factor_0.03discount.rate

forcings6$net.present.value_marginal.damages_additional.ton.carbon_0.025discount.rate = 
     forcings6$marginal.damages_additional.ton.carbon_dollars*forcings6$discount.factor_0.025discount.rate
forcings6$net.present.value_marginal.damages_additional.ton.CO2_0.025discount.rate = 
     forcings6$marginal.damages_additional.ton.CO2_dollars*forcings6$discount.factor_0.025discount.rate

forcings6$net.present.value_marginal.damages_additional.ton.carbon_0.05discount.rate = 
     forcings6$marginal.damages_additional.ton.carbon_dollars*forcings6$discount.factor_0.05discount.rate
forcings6$net.present.value_marginal.damages_additional.ton.CO2_0.05discount.rate = 
     forcings6$marginal.damages_additional.ton.CO2_dollars*forcings6$discount.factor_0.05discount.rate

```

Finally, we can compute the social cost of carbon for each of the three discount rates considered, for both carbon and CO2:

```{r}

social.cost.of.carbon = 
      data.frame(discount.rate = c(0.025,0.03,0.05),
                 cumulative.marginal.damages_additional.ton.carbon_dollars = 
                   c(sum(forcings6$net.present.value_marginal.damages_additional.ton.carbon_0.025discount.rate),
                     sum(forcings6$net.present.value_marginal.damages_additional.ton.carbon_0.03discount.rate),
                     sum(forcings6$net.present.value_marginal.damages_additional.ton.carbon_0.05discount.rate)),
                cumulative.marginal.damages_additional.ton.CO2_dollars = 
                c(sum(forcings6$net.present.value_marginal.damages_additional.ton.CO2_0.025discount.rate),
                     sum(forcings6$net.present.value_marginal.damages_additional.ton.CO2_0.03discount.rate),
                     sum(forcings6$net.present.value_marginal.damages_additional.ton.CO2_0.05discount.rate))  )

print(social.cost.of.carbon)


```