---
title: "CCE_Lab7"
author: "Ryan Peckner"
date: "8/10/2021"
output: html_document
---

Let's read in this week's forcings.

```{r}

options(digits = 15)

source("CCE_Lab1_MainFunctions.R")
library(readxl)

lab7_forcings = read_xlsx("CCE - Assignment 7 - Forcings.xlsx")
lab7_data = type.convert(as.matrix(lab7_forcings[c(3,8:13,17:19),-1]),as.is=FALSE)
lab7_data = t(lab7_data)
colnames(lab7_data) = make.names(lab7_forcings$Name[c(3,8:13,17:19)])
forcings7 = as.data.frame(lab7_data)
names(forcings7)[1] = "Year"
head(forcings7)

```

Now let's add consumption as a variable for each region, and use this to compute per capita consumption for each region.

```{r}

forcings7$Rich.consumption.trillions = 
      0.78*forcings7$Rich...Output...trill.
forcings7$Middle.consumption.trillions = 
      0.78*forcings7$Middle...Output...trill.
forcings7$Poor.consumption.trillions = 
      0.78*forcings7$Poor...Output...trill.

forcings7$Rich.per.capita.consumption.dollars = 
      (10^12)*forcings7$Rich.consumption.trillions/((10^6)*forcings7$Rich...Population..million.)
forcings7$Middle.per.capita.consumption.dollars = 
      (10^12)*forcings7$Middle.consumption.trillions/((10^6)*forcings7$Middle...Population..million.)
forcings7$Poor.per.capita.consumption.dollars = 
      (10^12)*forcings7$Poor.consumption.trillions/((10^6)*forcings7$Poor...Population..million.)

```

Now let's compute region-wise and global emissions.

```{r,cache=TRUE}

forcings7$Rich.emissions.MtC = 
  forcings7$Rich...emission.intensity.of.GDP..Mt.C..trill.*forcings7$Rich...Output...trill.
forcings7$Middle.emissions.MtC = 
  forcings7$Middle...emission.intensity.of.GDP..Mt.C..trill.*forcings7$Middle...Output...trill.
forcings7$Poor.emissions.MtC = 
  forcings7$Poor...emission.intensity.of.GDP..Mt.C..trill.*forcings7$Poor...Output...trill.

forcings7$Global.emissions.MtC = 
  forcings7$Rich.emissions.MtC + forcings7$Middle.emissions.MtC + forcings7$Poor.emissions.MtC

forcings7$atmospheric.carbon = sapply(1:nrow(forcings7),AtmosphericCO2,
                                      emissions=forcings7$Global.emissions.MtC)
forcings7$temp.increase = 
  sapply(1:nrow(forcings7),
         GlobalAvgTempIncrease_FromEmissions,
         emissions = forcings7$Global.emissions.MtC)

```

Let's now write down the modified damages function:

```{r}

phi_rich = 0.0546
phi_middle = 0.0505
phi_poor = 0.0443

psi_rich = -0.0013
psi_middle = -0.0017
psi_poor = -0.0015

Damages_Quadratic = function(temp.increase,phi,psi) {
  return(phi*temp.increase + psi*(temp.increase^2))
}

forcings7$Rich.damages = sapply(forcings7$temp.increase,
                                Damages_Quadratic,
                                phi=phi_rich,
                                psi=psi_rich)
forcings7$Middle.damages = sapply(forcings7$temp.increase,
                                Damages_Quadratic,
                                phi=phi_middle,
                                psi=psi_middle)
forcings7$Poor.damages = sapply(forcings7$temp.increase,
                                Damages_Quadratic,
                                phi=phi_poor,
                                psi=psi_poor)

forcings7$Rich.damages.trillions = forcings7$Rich.damages*forcings7$Rich...Output...trill.
forcings7$Middle.damages.trillions = forcings7$Middle.damages*forcings7$Middle...Output...trill.
forcings7$Poor.damages.trillions = forcings7$Poor.damages*forcings7$Poor...Output...trill.

```

Now we will compute marginal damages for each region.

```{r,cache=TRUE}

forcings7$extra.ton.emissions = forcings7$Global.emissions.MtC 
forcings7$extra.ton.emissions[2] = forcings7$extra.ton.emissions[2] + 1e-6

forcings7$temp.increase_extra.ton.emissions = 
  sapply(1:nrow(forcings7),
         GlobalAvgTempIncrease_FromEmissions,
         forcings7$extra.ton.emissions)

forcings7$Rich.damages_extra.ton.emissions = sapply(forcings7$temp.increase_extra.ton.emissions,
                                Damages_Quadratic,
                                phi=phi_rich,
                                psi=psi_rich)
forcings7$Middle.damages_extra.ton.emissions = sapply(forcings7$temp.increase_extra.ton.emissions,
                                Damages_Quadratic,
                                phi=phi_middle,
                                psi=psi_middle)
forcings7$Poor.damages_extra.ton.emissions = sapply(forcings7$temp.increase_extra.ton.emissions,
                                Damages_Quadratic,
                                phi=phi_poor,
                                psi=psi_poor)

forcings7$Rich.damages_extra.ton.emissions_trillions = forcings7$Rich.damages_extra.ton.emissions*forcings7$Rich...Output...trill.
forcings7$Middle.damages_extra.ton.emissions_trillions = forcings7$Middle.damages_extra.ton.emissions*forcings7$Middle...Output...trill.
forcings7$Poor.damages_extra.ton.emissions_trillions = forcings7$Poor.damages_extra.ton.emissions*forcings7$Poor...Output...trill.

forcings7$Rich.marginal.damages = 
  (10^12)*(forcings7$Rich.damages_extra.ton.emissions_trillions - 
             forcings7$Rich.damages.trillions)
forcings7$Middle.marginal.damages = 
  (10^12)*(forcings7$Middle.damages_extra.ton.emissions_trillions - 
             forcings7$Middle.damages.trillions)
forcings7$Poor.marginal.damages = 
  (10^12)*(forcings7$Poor.damages_extra.ton.emissions_trillions - 
             forcings7$Poor.damages.trillions)

```

Now we convert marginal damages into utility loss:

```{r}

rho = 0.01

forcings7$Rich_conversion.factor = 
  (1/forcings7$Rich.per.capita.consumption.dollars)*((1/(1+rho))^seq(1,nrow(forcings7)))
forcings7$Middle_conversion.factor = 
  (1/forcings7$Middle.per.capita.consumption.dollars)*((1/(1+rho))^seq(1,nrow(forcings7)))
forcings7$Poor_conversion.factor = 
  (1/forcings7$Poor.per.capita.consumption.dollars)*((1/(1+rho))^seq(1,nrow(forcings7)))


forcings7$Rich_utility.loss = 
  forcings7$Rich.marginal.damages*forcings7$Rich_conversion.factor
forcings7$Middle_utility.loss = 
  forcings7$Middle.marginal.damages*forcings7$Middle_conversion.factor
forcings7$Poor_utility.loss = 
  forcings7$Poor.marginal.damages*forcings7$Poor_conversion.factor

```

Finally, we compute the total damages over all times and all regions:

```{r}

SCC_utils = sum(forcings7$Rich_utility.loss +
                  forcings7$Middle_utility.loss + 
                  forcings7$Poor_utility.loss)
print(SCC_utils)

```
Finally, we convert this figure in utils back to dollars:

```{r}

scc = data.frame(Region=c("Rich","Middle","Poor"),
                 scc_dollars.per.tC = 
                   SCC_utils*c(forcings7$Rich.per.capita.consumption.dollars[1],
                               forcings7$Middle.per.capita.consumption.dollars[1],
                               forcings7$Poor.per.capita.consumption.dollars[1]))
scc$scc_dollars.per.tCO2 = (12/44)*scc$scc_dollars.per.tC
print(scc)

```

